
In this section, we will further evaluate the metrics to understand their distributions and relationship to housing units.

## Scatter Plots

Let's better understand some main varaible's relationships to housing units. Specifically, (1) Permits, (2) Evictions, (3) Change in Percent of Units  that is Owner-Occupied, and (3) Percent White.

```{r plot_predict_scatplot}

vars.ivs.disc = var.ivs.all[grepl(".pct.", var.ivs.all)]
vars.ivs.cont = var.ivs.all[!grepl(".pct.", var.ivs.all)]

fig_num = 2

var.dvs.plot = c(
  'units.tot.15_19',
  'permits.tot.15_19'
  )

dependent_variable = 'units.tot.15_19'
dv_lab = "Units"

plot_colors= c(
  "blue", "green", "red", "orange"
)

n_max = function(vector, n){
  new_vector = sort(vector, decreasing=TRUE)
  return(max(tail(new_vector, -1*(n))))
}
#n_max(c(5,4,69,100,3,2,1),2)

second_max = function(vector){
  return(n_max(vector,1))
}
second_max(c(5,4,69,100,3,2,1))

ylim_num = max(sf.predict[[dependent_variable]])
ylim = max(sf.predict[[dependent_variable]])# 
ylim = n_max(sf.predict[[dependent_variable]], 2)

scat_vars = c('permits.tot.15_19', 'evictions.tot.10_19', 'occ_own_ch.pct.90_18','wht.pct.10_15')
scat_labs = c("Permits 2015-19", "Evictions 2010-19", "Change in Pct Owned 1990-2018", 'Pct White 2010-15')

plot_scats = function(variable_number){
  plot_color = plot_colors[variable_number]
  variable_name = scat_vars[variable_number]
  variable_label = scat_labs[variable_number]
  fm_equation = paste(dependent_variable, "~", variable_name, sep="")
  
  fm = as.formula(fm_equation)
  units.tot.15_19_variable = lm(fm, data = sf.predict)
  coefficient = 
    round(
      units.tot.15_19_variable$coefficients[variable_name][1], 2)
  
  xlim = n_max(sf.predict[[variable_name]], 2)
  
  scat_plot = 
    ggplot(
      data = sf.predict,
      aes(
        x = sf.predict[[variable_name]],
        y = sf.predict[[dependent_variable]])) +
    geom_point(size=2, shape=20) +
    labs(title = 
           glue("Figure {fig_num}.{variable_number}: {variable_label}"),
         subtitle = glue("{fm_equation}     Coefficient = {coefficient %>% count_format()}")
         ) +
    geom_smooth(method = "lm", se=F,
                color = plot_color) +
    xlab(variable_name) +
    ylab(dv_lab) +
    xlim(min(sf.predict[[variable_name]]), xlim) + 
    ylim(min(sf.predict[[dependent_variable]]), ylim) + 
    plotTheme()
  print(scat_plot)}
```

The first plot shows housing units as they relate to permits. The coefficient of the two variables suggests that there is a positive relationship. More permits, more units. This makes sense. Another interesting pattern is that some of the block groups with the most units, don't necessarily have the highest amount of permits. Suggesting that there are larger projects that generate more units. 


```{r scat_plot1}
plot_scats(1)
```
The next plot highlights housing units over evictions. The coefficient suggests a positive relationship, but it is close. Many block groups that added alot of units didn't have alot of evictions. This could suggest that those with alot of new units had more evictions previously. But based on the maps in the last sections, there are alot of evictions spread throughout the city.  

```{r scat_plot2}
plot_scats(2)
```
Figure 2.3 shows the relationship in the "Change of Percent Owned from 1990 to 2018" and New Housing Units. It appears to be a positive coefficient. The more units added, the larger share in owner-occupied units. This could suggest that the new units added are likely condos. This is concerning if the share of apartment units are decreasing do to these new condos. 

```{r scat_plot3}
plot_scats(3)
```
Figure 2.4 shows the change in pct white from 2010 to 2015. In the related map in the previous section, it was visually apparent that the block groups with a growth in White people also saw more housing units added. This scatterplot suggests a similar conclusion. Although many block groups saw a loss in white people, the ones with added units saw larger increases. 

```{r scat_plot4}
plot_scats(4)
```

## Correlation
*Correlation Matrices, Independence Test*

Before performing any regression, it will be important to understand the linear relationship between (1) the predictors & themselves, and (2) the predictors and housing units. To understand this, we will use a Pearson Correlation matrix to weed out weak predictors and any predictors that are too related to one another. 

``` {r corr_cont_pearson, fig.width=8, fig.height=8, include = FALSE}

fig_num = 3


ggcorr_full = function(focus_df, #focus_vars= vars, 
                       focus_var = var.dv, corrmethod = 'pearson',
                       title=NULL, subtitle=NULL){
  num = ncol(sf.cont)
  
  BC = 'white'
  FC = "grey80"
  
  BL = 2
  FL = .2
  RL = 1:num-0.5
  RL[1] = .1
  ## RL[num] = .1
  
  library(reshape2)
  
  focus_df.melt = 
    cor(focus_df, method=corrmethod) %>%
    melt()
  
  
  xfaces = focus_df.melt$Var1 %>% as.character() %>% unique()
  yfaces = focus_df.melt$Var2 %>% as.character() %>% unique()
  
  change_focus = function(
    focus_vector, focus, 
    focus_change='bold', unfocus_change='plain'){
    vector = focus_vector[]
    vector[!(vector %in% c(focus))] = unfocus_change
    vector[vector %in% c(focus)] = focus_change
    return(vector)
  }
  
  xfaces = change_focus(xfaces, focus_var)
  yfaces = change_focus(yfaces, focus_var)
  
  focus_df.melt$value = 
    focus_df.melt$value %>%
    round(2)
  
  p = ggplot(focus_df.melt, aes(Var1, Var2, fill = value)) +
    geom_tile(color = "white") +
    geom_text(
      aes(Var1, Var2, 
          label = gsub("0\\.", "\\.", value)), 
      color = "black", size = 4) +
    scale_fill_gradient2(
      ## grey95
      low = "#6D9EC1", mid = "white", high = "#E46726",
      midpoint = 0, limit = c(-1,1), space = "Lab", 
                         name="Pearson\nCorrelation") +
    theme_minimal() +
    coord_fixed() +
    theme(
          axis.text.x = 
            element_text(
              face = xfaces,
              angle=45, vjust=1, hjust=1, 
              margin=margin(-2,0,0,0)),
          ## axis.text.x = element_text(margin=margin(-2,0,0,0)),  
          ###  Order: top, right, bottom, left  
          axis.text.y = 
            element_text(
              face = yfaces,
              margin=margin(0,-2,0,0)),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          legend.position="none") +
    labs(title = title, subtitle=subtitle) + 
    geom_vline(xintercept=RL, colour=BC, size=BL) +
    geom_hline(yintercept=RL, colour=BC, size=BL) +
    geom_vline(xintercept=RL, colour=FC, size=FL) +
    geom_hline(yintercept=RL, colour=FC, size=FL)
  return(p)
  }
```

The correlation plots are split in two so that all of the metrics can be viewed. I also already removed alot of metrics that were too similar. We want to prevent multicollinearity and weak predictors.

``` {r plot_corr, fig.width=10, fig.height=10}


remove_col = c(
  'hinc.tot.18',
  #'units.tot.00_14',
  'units.tot.00_04',
  'units.tot.05_09',
  'units.tot.10_14',
  'permits.tot.00_04',
  'permits.tot.05_09',
  'permits.tot.10_14',
  'hh.tot.18',
  "permits.tot.00_14",
  #"breach_lease.tot.10_19",
  #"change_use.tot.10_19",
  "reno.tot.10_19",
  "renter_payment.tot.10_19",
  "evictions.tot.10_19",
  #"evictions.tot.00_19"
  "pop.tot.18",
  'white.tot.00',
  'hh.tot.90',
  'mov.pct.12',
  'inc.med.12',
  'hh_costs.med.18',
  'hh_inc.med.90',
  'mhval_ch.tot.12_18',
  'occ_own.pct.90',
  'occ_own.pct.18',
  'occ_rent.pct.18',
  'FAR.avg.2019',
  'pop.tot.00',
  'hh_inc.med.00',
  'occ_rent_ch.pct.90_18',
  'occ_own_ch.pct.90_18',
  'occ_rent.tot.18',
  'occ_rent_ch.pct.00_18',
  'lowinc.pct.00',
  'area_res.pct.2019',
  'units.tot.00',
  'income.med.18',
  'rent.med.90',
  'hval.med.00',
  'rent.med.00',
  'hh.tot.00',
  'rent.med.90',
  'mhval.tot.18',
  'white.tot.90',
  'hval.med.90',
  'pop.tot.90',
  'lotarea.avg.2019',
  'occ_own_price.med.12',
  'occ_rent_price.med.12',
  'mrent.tot.18',
  'dist_BART.tot.2020',
  'res_area',
  'units_p_parcel.avg.2019',
  'units_p_parcel_r.avg.2019',
  'lotarea_res.tot.2019',
  'lotarea.avg.2019',
  'bldgsqft.avg.2019',
  'white.tot.90',
  'rent.med.00',
  'mrent.tot.18',
  'mhval.tot.18',
  'occ_own.pct.00',
  #'college_ch.pct.00_18',
  'lowinc.pct.90'
  # 'breach_lease.tot.10_19',
  # 'white.tot.18',
  # 'mhval_ch.pct.00_18',
  # 'poc_linc.pct.10_15',
  # 'hh_inc.med.18',
  # 'mhval_ch.pct.00_18',
  # 'pop25.tot.18',
  # 'units.avg.2019',
  # 'area'
  )

vars.ivs.cont_fin = vars.ivs.cont[!(vars.ivs.cont %in% remove_col)]


vars.ivs.disc_fin = vars.ivs.disc[!(vars.ivs.disc %in% remove_col)]


len = length(vars.ivs.cont_fin)
len_half = round(len/2)

var.dvs.plot = c(
  'units.tot.15_19',
  'permits.tot.15_19'
  )

sf.cont = 
  sf.predict %>%
  select(c(var.dvs.plot, vars.ivs.cont_fin)) %>% 
  st_drop_geometry(.)

plot = ggcorr_full(sf.cont,
            title = glue("Figure {fig_num}: Pearson Correlation"),
            subtitle = "Continous Variables"
)
print(plot)

```
From this first correlation plot, we can see some interesting relationships. Permit & Unit counts aren't necessarily the most positive relationship -- especially considering the scatter plots. Instead the change in white people is the highest correlation (wht.pct.10_15). Then the pct of people who have recently moved (mov.pct.18). An interesting negative correlation is the percent of building square footage that is residential (bsqft_res.pct.2019). An interesting null correlationship is the percent of people who have a college degree. I would imagine it is due to their being college educated people who own houses in the suburban outskirts of SF as well as the downtown. 


```{r pear_cont, fig.width=6, fig.height=6}
sf.cont = 
  sf.predict %>%
  select(c(var.dvs.plot, vars.ivs.disc_fin[!(vars.ivs.disc_fin %in% vars.ivs.cont_fin)])) %>% 
  st_drop_geometry(.)


plot = ggcorr_full(sf.cont,
            title = glue("Figure {fig_num}: Pearson Correlation"),
            subtitle = "Continous Variables"
)
print(plot)
```
The second correlation plot has more lower powered correlations. None of the predictors stick out.

The next section will look more into the clustering effect of housing and permits. 